
I would like to express sincere gratitude to my dissertation advisor, Professor Abani K. Patra, for his guidance, solid support, encouragement, and constant proposal of challenging new ideas. I am grateful to him for all trainings that I got in his group and the academic freedom while doing research. The opportunities I had during my dissertation work to attend conferences are deeply appreciated, too.
I also express my appreciation to all the other members of my supervisory committee, Professor E. Bruce Pitman, Professor Marcus Bursik, and Professor Paul Bauman. In particular, the advice and lecture-style instructions from Professor E. Bruce Pitman has been of great help in model establishment, developing and implementing new SPH schemes based on solving local Riemann problems, such as GSPH and RSPH. Discussions with Professor Marcus Bursik regarding the necessity of developing a new plume model, the volcanic model simplification and potential implementations of the model are greatly helpful and appreciated. I would also like to take the chance to thank him for his availability and his help to keep me on track. Advice and encouragement from Professor Bauman are greatly appreciated. I am also thankful for his time and efforts on reading my dissertation and giving constructive comments. 

I am grateful to collaborators of Professor Patra as well. Particularly, Dr. Jones provided a lot of helps in dealing with technical issues at the software development stage and implementation stage. I always got quick, patient and professional response from him even for small technical questions. Great appreciation goes to him. Advices and helps from Dinesh, Ramona, Hossein and Topher made it easier for me to get start in the lab. I have been discussing and shearing knowledge and skills with Ali, Prashant, Qingyuan Yang, Abishek, Renette and Lekha, which were pleasant learning experience and helpful to this dissertation. Data shared by peers are indispensable for making progress during this dissertation work, great appreciation goes to them: Suzuki Yujiro, Tomaso Esposti Ongaro and Antonio Costa.

Finally, I express my deep gratitude to my family for their love and support.